\Huge \textbf{Hybrid rendering}\\ \huge using Radeon Rays

### 2018-11-03 Paper/Paper_BillietStijn_RadeonRays.tex
Worked on paper (primarily parts about gbuffers and layout)

\subsection{Raytracing}
\subsubsection{OpenCL}
Open Computing Language is a framework for writing programs that execute across compute devices. This includes among other hardware accelerators, the central processing unit (CPU) and the graphics processing unit (GPU). These aforementioned programs are called kernels and are also written on a pseudo c-language. OpenCL provides the programmer of a standard interface to perform parallel computing tasks. The programmer has to set up the OpenCL execution pipeline and write the kernels. Under this section you'll find a quick rundown of aforementioned execution 
model.\newline

@@

\begin{itemize}
	\item{Platform}\newline
	The platform that we are running this code on, so which vendor (e.g. Nvidia).
	\item{Device}\newline
	What device are we going to run this on could be anything from CPU, GPU to APU.
	\item{Context}\newline
	This is the runtime interface between the host program (run on the CPU) and the device (CPU/GPU/etc.). The context is in charge of managing all the OpenCL resources so programs, kernels, command queues and buffers.
	\item{Program}\newline
	The entire OpenCL program, this could be one or more kernels/device functions bundled together into a single program file.
	\item{Kernel}\newline
	The starting point/building blocks of a OpenCL program. Kernels are called from the CPU, they provide the basic units of executable code that run on a OpenCL device. The kernels are preceded with the keyword "\_\_kernel"
	\item{Command queue}\newline
	The command queue allows the kernels execution commands to be sent to the device. This effectively queues them for execution, this en-queuing can be in-order or out-of-order.
	\item{Memory objects}\newline
	The memory objects are what is provided to the kernels. Kernels do accept basic types like floats, ints, bools etc. and self defined types like Ray, Intersection etc, but if there is a large buffer of those it's kept in a Memory buffer (e.g. images).
\end{itemize}

@@

\begin{itemize}
	\item{Global memory}\newline
	Global memory is very similar to RAM, it is the largest memory buffer but also the slowest. Global memory can be written to / read from by all work items (or threads).
	\item{Constant memory}\newline
	A small chunk of global memory on the device, can be read from by all work items, but can not be written to. However the host application has full read/write access to this memory pool. Constant memory is a bit faster than 
	\item{Local memory}\newline
	The local memory is a shared memory pool among work items in the same work group, similar to cache on the CPU. Local memory allows work items to share their results with work items belonging in the same work group. Local memory is up to 100 times faster than global memory.
	\item{Private memory}\newline
	The fastest type of memory, similar to registers on the cpu. Each work item (thread) has a tiny buffer of private memory to store intermediate results in.
\end{itemize}

@@

\subsection{Rasterization}
\subsubsection{OpenGL}
\begin{itemize}
	\item{OpenGL}
	\pagebreak
	\item{Loading libraries}\newline\par
	OpenGL is a specification that has many different versions, implementations and features. The hardware manufacturers have to make sure their hardware/drivers provides the graphics programmers of those features. However when working with immensely different target platforms, ranging from old hardware to new and from obscure cards to bleeding edge ones the graphics programmers need some kind of way to determine what's possible on that specific piece of hardware. That is where loading libraries come in, loading libraries essentially provide the programmer with function pointers to OpenGL agnostic functions at runtime, meaning they provide us the memory addresses of those functions seeing they depend from hardware to hardware.\newline\par

@@

	There are numerous libraries out there that do just that, libraries like GLEW, GL3W, Glad, Glee and many more. All libraries that exist out there all deliver on the same promise, meaning that the choice of library really boils down to personal preference.\newline\par
	\item{GLM}\newline\par
	When it comes to math libraries that plug in nicely with an OpenGL renderer there's really only one good option namely GLM. OpenGL Mathematics (GLM) is a header only C++ mathematics library for graphics software based on the OpenGL Shading Language (GLSL). GLM exposes a well defined interface for everything from general calculus to quaternion rotations (e.g. vector operations, matrix calculations).
\end{itemize}
\subsubsection{Window creation}
A window allows the rendering specification to offload its framebuffer to the viewport, it also acts as a layer between the operating system and the game when it comes to input etc. The creation of a window is entirely dependent on the operating system it's creating the window for. For example when creating a window on linux the X window system should be addressed, whereas on Windows you would need to hook on to the WinMain function and use the WndProc callback to fetch any input. Therefore it is much smarter to use a library that hides all this operating system agnostic stuff in an abstraction layer.\newline\par
SDL, SFML and GLFW are all low level libraries that handle multimedia indirection. They all deliver the same features in some shape or form. The choice really boils down to maturity of the library, documentation, usefulness of its features and frequency of the updates (if any).
\subsubsection{Model loading}
\subsubsection{Texture loading}
\pagebreak

@@

The third rendering pipeline being forward+ although the most performant will not be covered with explicit detail. The reason that this technique is not that applicable has primarily to do with the fact that we have no way of postponing lighting effects. This technique also covers lighting object per object and that is precisely the opposite of what raytracing is trying to overcome.\newline\par
Forward+ is a rendering technique that improves upon regular forward rendering. Forward+ first determines which lights are overlapping which ares of the screen (screen space). During the shading phase that list is then consulted to only account for the lights that are potentially overlapping the current pixel fragment. The forward+ technique consists primarily of three simple passes.\newline\par

@@

\subsubsection{Ray construction using g-buffers}

@@

\subsection{Related work}
\pagebreak

@@

\subsection{Coordinate systems}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam. Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus. Sed vel ligula justo, sit amet vestibulum dolor. Sed vitae augue sit amet magna ullamcorper suscipit. Quisque dictum ipsum a sapien egestas facilisis.

\subsection{Model loading}
Assimp as the name implies is an asset import library that handles about every relevant model export type that exist. The assimp api needs to be built using CMake and can then be linked into our solution. Then it's just a simple case of creating an instance of their importer context and calling the readfile method.
\subsection{Texture loading}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam. Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus. Sed vel ligula justo, sit amet vestibulum dolor. Sed vitae augue sit amet magna ullamcorper suscipit. Quisque dictum ipsum a sapien egestas facilisis.
\subsection{Deferred shading}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam. Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus. Sed vel ligula justo, sit amet vestibulum dolor. Sed vitae augue sit amet magna ullamcorper suscipit. Quisque dictum ipsum a sapien egestas facilisis.
\section{Implementation ray-tracing}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam. Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus. Sed vel ligula justo, sit amet vestibulum dolor. Sed vitae augue sit amet magna ullamcorper suscipit. Quisque dictum ipsum a sapien egestas facilisis.
\subsection{OpenCL implementation}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam. Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus. Sed vel ligula justo, sit amet vestibulum dolor. Sed vitae augue sit amet magna ullamcorper suscipit. Quisque dictum ipsum a sapien egestas facilisis.

\subsection{Radeon rays implementation}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam. Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus. Sed vel ligula justo, sit amet vestibulum dolor. Sed vitae augue sit amet magna ullamcorper suscipit. Quisque dictum ipsum a sapien egestas facilisis.
\subsection{Shadow implementation}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam. Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus. Sed vel ligula justo, sit amet vestibulum dolor. Sed vitae augue sit amet magna ullamcorper suscipit. Quisque dictum ipsum a sapien egestas facilisis.
\subsection{Optimizations}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis risus ante, auctor et pulvinar non, posuere ac lacus. Praesent egestas nisi id metus rhoncus ac lobortis sem hendrerit. Etiam et sapien eget lectus interdum posuere sit amet ac urna. Aliquam pellentesque imperdiet erat, eget consectetur felis malesuada quis. Pellentesque sollicitudin, odio sed dapibus eleifend, magna sem luctus turpis, id aliquam felis dolor eu diam. Etiam ullamcorper, nunc a accumsan adipiscing, turpis odio bibendum erat, id convallis magna eros nec metus. Sed vel ligula justo, sit amet vestibulum dolor. Sed vitae augue sit amet magna ullamcorper suscipit. Quisque dictum ipsum a sapien egestas facilisis.

\chapter{Conlusion}
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.

>>>>

\subsection{Raytracing}
\subsubsection{OpenCL}
Open Computing Language is a framework for writing programs that execute across compute devices. This includes among other hardware accelerators, the central processing unit (CPU) and the graphics processing unit (GPU). These aforementioned programs are called kernels and are also written on a pseudo c-language. OpenCL provides the programmer of a standard interface to perform parallel computing tasks. The programmer has to set up the OpenCL execution pipeline and write the kernels.

@@

Under this section you'll find a quick rundown of aforementioned execution 
model.\newline\par
\begin{itemize}
	\item{\textbf{Platform}}\newline\par
	The platform that we are running this code on, so which vendor (e.g. Nvidia).
	\item{\textbf{Device}}\newline\par
	What device are we going to run this on could be anything from CPU, GPU to APU.
	\item{\textbf{Context}}\newline\par
	This is the runtime interface between the host program (run on the CPU) and the device (CPU/GPU/etc.). The context is in charge of managing all the OpenCL resources so programs, kernels, command queues and buffers.
	\item{\textbf{Program}}\newline\par
	The entire OpenCL program, this could be one or more kernels/device functions bundled together into a single program file.
	\item{\textbf{Kernel}}\newline\par
	The starting point/building blocks of a OpenCL program. Kernels are called from the CPU, they provide the basic units of executable code that run on a OpenCL device. The kernels are preceded with the keyword "\_\_kernel"
	\item{\textbf{Command queue}}\newline\par
	The command queue allows the kernels execution commands to be sent to the device. This effectively queues them for execution, this en-queuing can be in-order or out-of-order.
	\item{\textbf{Memory objects}}\newline\par
	The memory objects are what is provided to the kernels. Kernels do accept basic types like floats, ints, bools etc. and self defined types like Ray, Intersection etc, but if there is a large buffer of those it's kept in a Memory buffer (e.g. images).
\end{itemize}

@@

\begin{itemize}
	\item{\textbf{Global memory}}\newline\par
	Global memory is very similar to RAM, it is the largest memory buffer but also the slowest. Global memory can be written to / read from by all work items (or threads).
	\item{\textbf{Constant memory}}\newline\par
	A small chunk of global memory on the device, can be read from by all work items, but can not be written to. However the host application has full read/write access to this memory pool. Constant memory is a bit faster than 
	\item{\textbf{Local memory}}\newline\par
	The local memory is a shared memory pool among work items in the same work group, similar to cache on the CPU. Local memory allows work items to share their results with work items belonging in the same work group. Local memory is up to 100 times faster than global memory.
	\item{\textbf{Private memory}}\newline\par
	The fastest type of memory, similar to registers on the cpu. Each work item (thread) has a tiny buffer of private memory to store intermediate results in.
\end{itemize}

@@

\subsection{Rasterization}
\subsubsection{OpenGL}
\begin{itemize}
	\item{\textbf{OpenGL}}
	\pagebreak
	\item{\textbf{Loading libraries}}\newline\par
	OpenGL is a specification that has many different versions, implementations and features. The hardware manufacturers have to make sure their hardware/drivers provides the graphics programmers of those features. However when working with immensely different target platforms, ranging from old hardware to new and from obscure cards to bleeding edge ones the graphics programmers need some kind of way to determine what's possible on that specific piece of hardware. That is where loading libraries come in, loading libraries essentially provide the programmer with function pointers to OpenGL agnostic functions at runtime, meaning they provide us the memory addresses of those functions seeing they depend from hardware to hardware.\newline\par

@@

	There are numerous libraries out there that do just that, libraries like GLEW, GL3W, Glad, Glee and many more. All libraries that exist out there all deliver on the same promise, meaning that the choice of library really boils down to personal preference.\newline\par
	\item{\textbf{GLM}}\newline\par
	When it comes to math libraries that plug in nicely with an OpenGL renderer there's really only one good option namely GLM. OpenGL Mathematics (GLM) is a header only C++ mathematics library for graphics software based on the OpenGL Shading Language (GLSL). GLM exposes a well defined interface for everything from general calculus to quaternion rotations (e.g. vector operations, matrix calculations).
\end{itemize}
\subsubsection{Window creation}
A window allows the rendering specification to offload its framebuffer to the viewport, it also acts as a layer between the operating system and the game when it comes to input etc. The creation of a window is entirely dependent on the operating system it's creating the window for. For example when creating a window on linux the X window system should be addressed, whereas on Windows you would need to hook on to the WinMain function and use the WndProc callback to fetch any input. Therefore it is much smarter to use a library that hides all this operating system agnostic stuff in an abstraction layer.\newline\par
SDL, SFML and GLFW are all low level libraries that handle multimedia indirection. They all deliver the same features in some shape or form. The choice really boils down to maturity of the library, documentation, usefulness of its features and frequency of the updates (if any).
\subsubsection{Texture loading}
A texture is a 2D (could be 3D or 1D) image packed with color data that is used to feed into the graphics pipeline to texturize geometry. Textures can be stored in a dozen of file formats, each with their own structure to store the pixel data. Textures are stored in a binary format and can therefore be parsed by a binary reader. The problem however with doing textureloading is that every format is different so the programmer would be in charge of writing a dozen different loaders to support the most basic of texture types. Seeing that this is rather cumbersome most people would opt for a library that does this for them. Again there are quite a lot of libraries out there that do this. \newline\par

SOIL, GLi, DevIL, FreeImage and stb\_image are all texture loading libraries that promise on loading the most prevalent of types.
\pagebreak

\subsubsection{Model loading}
\begin{itemize}
	\item{\textbf{3D file format}}\newline\par
	A 3D file format is a specification to store information about 3D models in the form of plain text or as binary data. The main purpose of such format is to encode 3D model's geometry, material information and animation data. Seeing there are loads of formats out there, creating your own model loading library has become something of a novelty. Fortunately there are loads of libraries out there that load about any format you throw at it.
	\item{\textbf{AssImp}}\newline

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.9\textwidth]{assimpstructure.png}
	\end{figure}

	Asset import or AssImp for short is a model loading library designed to load as many formats as possible and to unify them into a searchable structure to be used by the host application. It structures the entire scene into a scene object, which consecutively holds a rootnode and a collection of meshes and materials. That rootnote then has leads to other nodes that all hold the same information namely a list of indices to the corresponding meshes and a next connection. This leads to a clear distinction as to what are seperate objects in the global model file, which can then later be used by the game-engine to ensure perfect relationships between aforementioned meshes.
\end{itemize}

@@

The third rendering pipeline, although the most performant will not be covered with explicit detail. The reason that this technique is not that applicable has primarily to do with the fact that we have no way of doing per world space lighting effects. Forward+ is a rendering technique that improves upon regular forward rendering. Forward+ first determines which lights are overlapping which areas of the screen (screen space). During the shading phase that list is then consulted to only account for the lights that are potentially overlapping the current pixel fragment. The forward+ technique consists primarily of three simple passes. In the light culling pass, each light in the scene is sorted into screen space tiles. In the opaque pass, the light list generated from the light culling pass is used to compute the lighting for opaque geometry. The transparent pass is similar to the opaque pass in the sense that the light list is again consulted when shading the transparent pixel fragments.

@@

\subsubsection{Raytracing using g-buffers}

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{gbuffer.png}
	\caption{Illustration of g-buffers}
\end{figure}

As touched earlier deferred rendering fills the off screen render buffers (g-buffers) with scene information in its first pass. This effectively enables us to do rendering after the fact by then using that information to construct everything we need to apply our lighting.\newline\par

The information that is stored in the g-buffers is entirely up to the programmer (e.g. position data, normal data, color data etc.). However to be able to use this in conjunction with ray construction the programmer needs atleast the position data and the normal data. Some people would however argue that the position data isn't necessary and that the position can be calculated by the z-buffer and that would be an entirely correct answer, but for simplicity sake people opt for the position data instead.\newline\par

With the geometry data now in the g-buffers, that data can now be used to construct the rays. The ray construction can be sped up by doing it on the GPU, seeing that this operation is highly parallel in nature. Essentially what it boils down to is to make sure that meshes have been synchronized with the raytracer, meaning that the raytracer has all the right meshes in it's scene to be able to fetch all occlusion/intersections. Then quite simply sample all the position entries and normal entries (for the entire screen) and send out a ray for every pixel, this will in turn correlate to the raytracer api seeing those position entries should correspond with the meshes in the scene.\newline\par

Once the rays have been constructed the raytracer can now simply query the occlusion/intersections with the scene geometry. That data can then be used to shade the object respectively.\newline\par

@@

@@

\subsection{Coordinate systems}

\subsection{Model loading}

\subsection{Texture loading}

\subsection{Deferred shading}

\section{Implementation ray-tracing}

\subsection{OpenCL implementation}

\subsection{Radeon rays implementation}

\subsection{Shadow implementation}

\subsection{Optimizations}

\chapter{Conclusion}
